---
title: "R-Dataframes"
output:
  pdf_document: default
  html_notebook: default
---


# Dataframes
In any form of data analysis, it is most important to understand the datasets. 
Data frames are a form of data structures in tabular format. They are most widely used for data analysis in R. 
Data frames,:

- rows : Observations
- columns are vectors of particular data types (such as character, integer,logical).
- different columns can be diff datatype.
- Elements of same column should be same type


Creating a data frame
```{r}
firstName <- c("Vilas","Pallavi", "Archana","Matti", "Dylan","Cheryl","Fahad") ## firstName is a Vector
lastName <- c("Menon", "Gaur", "Yadav", "Lam", "Lee","Pan","Paryani")
Gender <- c("male", "female", "female", "male", "male","female","male")
id_no <- c(20,29,32,15,26,19,30)
Fav_icecream <- c("Vanilla","Rose","Saffron","Chocolate","Vanilla","Chocolate","Vanilla")


Menon_lab <-  data.frame(firstName, lastName, FullName=paste(firstName, lastName), Gender, id_no, Fav_icecream) ## data.frame is a function, Menon_lab is dataframe
Menon_lab
```

```{r}
summary(Menon_lab)
```
```{r}
class(id_no)
```

```{r}
table(Menon_lab$Gender)

table(Menon_lab$Gender, Menon_lab$Fav_icecream)
```


```{r}
#save this table
write.csv(Menon_lab,"menon_lab.csv")
```

```{r}
# open and read a csv file
m_lab <- read.csv("lab.csv")
m_lab
```
```{r}
getwd()
```

```{r}
list.files()
```

```{r}
##Exploring data frame

#Size
#dim(Menon_lab) #returns number of rows and the number of columns

#nrow(Menon_lab) # number of rows

ncol(Menon_lab) # number of columns



```

```{r}
#Content
head(Menon_lab)
tail(Menon_lab)
colnames(Menon_lab)
rownames(Menon_lab)
```
```{r}

str(Menon_lab)

```

```{r}
summary(Menon_lab)
```


####Indexing and subsetting
use [] brackets, [row position,column position] 
```{r}
Menon_lab
```

```{r}
#extract entire row, which is a dataframe with single observation

Menon_lab[2,]
```


```{r}
#Extract entire column, which is a vector
Menon_lab$firstName
Menon_lab[,"firstName"]
Menon_lab[,2]
```

```{r}
Menon_lab
```


```{r}
#Extract particular cell
Menon_lab[4,1]
```
```{r}
#extract cell using column name
Menon_lab[4,"firstName"]
```


```{r}
# subsetting
#extract n number of rows
Menon_lab[c(1,2,4),]
```

```{r}
##extract some rows and columns
new_set<- Menon_lab[1:3,c("firstName","id_no")]
new_set
```


```{r}
#extract rows based on some comparison
Menon_lab[Menon_lab$Gender=="female",]
```


```{r}
##make a new table with only female members
Menon_lab[Menon_lab$Gender=="female",]
female_members <- Menon_lab[Menon_lab$Gender=="female",]
female_members

```


```{r}
#show everything except particular column
Menon_lab[,-1]
```

```{r}
#show everything except particular rows
Menon_lab[-c(1,3),]
```
###using Grep

```{r}
###using Grep
## grep index values i.e.tell me the name of row numbers where the last name starts with 'L'
grep("^L", Menon_lab$lastName) # ^ is a regular expression also called regex in coding language

```



```{r}
####subsetting with grep
Menon_lab[grep("^L",Menon_lab$lastName),]


```






